Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-07 paper/main.tex
wilkenson calls continuity topological properties

\begin{LaTeXdescription}
  \item [continuity] how elements in a dataset are connect to each other, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
  \item [equivariance] functions on data that have an equivalent effect on the graphical representation, e.g. rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
\end{LaTeXdescription}

\subsection{Continuity}
\begin{figure}[!h]
  \includegraphics[width=\columnwidth]{k_different_types.png}
  \caption{Continuity is how elements in a data set are connected to each other, which is distinct from how the data is structured. The rows in (a) are discrete, therefore they have discrete continuity as illustrated by the discrete dots. The gaussian in (b) is a 1D continuous function, therefore the continuity of the elements of the gaussian can be represented as a line on an interval (0,1). In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
  \label{fig:related-work:continuity}
\end{figure}
We propose that a robust model of continuity, which is how the elements in a dataset are connected to each other, provides a way to develop library components that \note{finish this sentence in a sensible way}

@@

The model described in this work provides a method for expressing different types of continuities and nested continuities using the same model. We obtain this generality by using the mathematical theory of fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}. In this paper, we build on their work that proposes using topological spaces to represent different properties of data.

Specifically, Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection of topological spaces. 
\begin{definition}
A topological space $(X, \mathscr{T})$ is a set $X$ with a topology $\mathscr{T}$. Topologies are collections of open sets $U$ such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}$ \note{finish out this definition}
\end{definition}

@@

\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
  \label{def:category:K} 
  \begin{itemize}
    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space\cite{bradleyTopologyCategoricalApproach2020, } $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
  \end{itemize} 
\end{definition}

@@

\subsection{Union of Artists}
\label{sec:artist:union}
\begin{figure}[!h]
\centering
\subfloat[Union of H]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
\label{fig_first_case}}
%\hfil
\subfloat[Artist with shared/harmonized \vsection]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
\label{fig_second_case}}
\caption{Coherance? }
\label{fig_sim}
\end{figure}
\note{do we define coherant visualizations as one with shared inclusions?}

>>>>

\begin{LaTeXdescription}
  \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
  \item [equivariance] functions on data that have an equivalent effect on the graphical representation, e.g. rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
\end{LaTeXdescription}

\subsection{Continuity}
\begin{figure}[!h]
  \includegraphics[width=\columnwidth]{k_different_types.png}
  \caption{Continuity is how elements in a data set are connected to each other, which is distinct from how the data is structured. The rows in (a) are discrete, therefore they have discrete continuity as illustrated by the discrete dots. The gaussian in (b) is a 1D continuous function, therefore the continuity of the elements of the gaussian can be represented as a line on an interval (0,1). In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
  \label{fig:related-work:continuity}
\end{figure}
Continuity, which Wilkinson terms topological properties \cite{wilkinsonGrammarGraphics2005}, describes how elements in a data set are organized. Wilkenson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region. We propose that a robust model of continuity, which is how the elements in a dataset are connected to each other, provides a way to develop library components that can

@@

The model described in this work provides a modol for expressing data sets with different topological properties. 
We obtain this generality by using the mathematical theory of fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}. In this paper, we build on their work that proposes using topological spaces to represent different properties of data.

Specifically,Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection of topological spaces. 
\begin{definition}
A topological space $(X, \mathscr{T})$ is a set $X$ with a topology $\mathscr{T}$. Topologies are collections of open sets $U$ such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}$ \note{finish out this definition}
\end{definition}

@@

\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
  \label{def:category:K} 
  \begin{itemize}
    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
  \end{itemize} 
\end{definition}

@@

\subsection{Union of Artists}
\label{sec:artist:union}
\begin{figure}[!h]
\centering
\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
\label{fig:artist:union:glue}}
%\hfil
\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
\label{fig:artist:union:continuity}}
\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal. The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image. \note{add a border/make the same size/etc so 1 doesn't run over into two}}
\end{figure}

Visualizations generally consist of more than one graphic, for example two line plots, a legend, axis ticks, spines, and labels. \note{formalize superposition, than break down what that means for composition}

These graphics can be generated by disjoint data inputs  $(\dtotal_i: i\in I)$ into different artist functions $\vartist$ with the same output target space $\gtotal$. We define this composition via the disjoint union of the inputs 
\begin{equation}
  \label{eq:artist:union:glue}
+ \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
\end{equation}
This approach does not guarantee a coherent image, as illustrated in \autoref{fig:artist:union:glue}, because each data is explicitly considered as distinct from the others.

Many times though, graphics are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}. In these situations, we consider each of the data inputs as component bundles over a shared continuity \dbase.

\begin{equation}
  \label{eq:math:data:base:decompose}
  \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
\end{equation}

The sections $\dsection$ on each component bundle $\dtotal_i$ over a point $\dbasepoint \in \dbase$ are the elements of the data associated with that point. We use this property to establish which of these elements are the same, because they come from the same fiber, and therefore should share transformation functions to visual elements. This bookkeeping yields graphics with consistent visual encoding of data elements, as shown in \autoref{fig:artist:union:continuity}.
